# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

* Parse raw data

```{r}
rawData <- read.csv("activity.csv")
```

* Aggregate sum of steps by date and mean of steps by interval accross dates

```{r}
stepsByDate <- aggregate(rawData$steps,by=list(Date=rawData$date),na.rm=TRUE,FUN=sum)
names(stepsByDate) <- c("Date","Steps")
head(stepsByDate)

stepsByInterval <- aggregate(rawData$steps,by=list(Date=rawData$interval),na.rm=TRUE,FUN=mean)
names(stepsByInterval) <- c("Interval","Steps")
head(stepsByInterval)
```

## What is mean total number of steps taken per day?

```{r}
hist(stepsByDate$Steps)
```

```{r}
mean(stepsByDate$Steps)
median(stepsByDate$Steps)
```


## What is the average daily activity pattern?

```{r}
plot(stepsByInterval$Interval, stepsByInterval$Steps, type="l")
```

* The maximum number of average steps per interval is:

```{r}
stepsByInterval[stepsByInterval$Steps == max(stepsByInterval$Steps),]$Interval
```


## Imputing missing values

* The number of NAs is:

```{r}
sum(is.na(rawData$steps))
```

* Strategy for NA replacement: substituting NAs by the mean of that day

```{r}
filledData <- rawData

for(i in 1:nrow(filledData)) {
  if(is.na(filledData$steps[i])) {
     filledData$steps[i] <- stepsByDate[stepsByDate$Date == filledData$date[i],]$Steps[1]
  }
}

fdSteps <- aggregate(filledData$steps,by=list(Date=filledData$interval),FUN=mean)
names(fdSteps) <- c("Interval","Steps")
head(fdSteps)
```

* Histogram, mean and median of the new data

```{r}
hist(fdSteps$Steps)
mean(fdSteps$Steps)
median(fdSteps$Steps)
```


## Are there differences in activity patterns between weekdays and weekends?

```{r}
stepsByDate$Day <- weekdays(as.Date(stepsByDate$Date))
stepsByDate$WD <- stepsByDate$Day
for(i in 1:nrow(stepsByDate)) {
  if(stepsByDate$Day[i] == "Saturday" || stepsByDate$Day[i] == "Sunday")
    stepsByDate$WD[i] <- "Weekend"
  else
    stepsByDate$WD[i] <- "Weekday"
}

head(stepsByDate)
```

* Plot steps by type of day

```{r}
library(lattice)
with(stepsByDate, xyplot(Steps ~ Date | WD, type="l", layout=c(1,2)))
```